Type Refinements for the Working Class

### 2015-12-19 paper.tex
Fix backwards yoneda ⇒ coyoneda

  So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of
  explaining such judgments in terms of the presheaf topos is that we do not
  need to deal with the complexities of quantifying over renamings, since this
  is implicit in the definition of the exponential of presheaves,
  \begin{align*}
    \ADefine{
      B^A(\Upsilon)
    }{
      \Hom{\Sets^\SCtx}{
        \Yoneda{\Upsilon}\times A
      }{
        B
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}
      (\Upsilon\inj\Upsilon')
      \Rightarrow
      {B(\Upsilon')}^{A(\Upsilon')}
    }
  \end{align*}

@@

%% In general, we must explain typehood for parametric refinements
%% $\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
%% Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
%% transformation
%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
%% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
%% the following diagram commutes, where
%% $\lambda(\chi)$ is the exponential transpose of $\chi$:
%% \[
%%   \IsValueType{\Upsilon}{\phi}{\tau}
%%   \triangleq
%%   \begin{tikzcd}[sep=huge]
%%     \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
%%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
%%       \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
%%     \IMode{\Omega}\\
%%     \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
%%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
%%   \end{tikzcd}
%% \]
%%
%% Equivalently, and perhaps more simply, we can express value typehood internally
%% in our topos as a presheaf:
%% \[
%%   \Define{
%%     \IsValueType{-}{\phi}{\tau}
%%   }{
%%     \sum_{\psi\in\Pow{\Values_\tau^2}}
%%     \prod_{M,N\in\Values_\tau}
%%     {\phi(M,N)}^{\psi(M,N)}
%%   }
%% \]

@@

Now, let us return briefly to Martin-L\"of's notion of type, which we call a
\emph{value type}. The minimal data of such a type's definition consists in a
partial equivalence relation on values,
$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
It is easy to turn such a relation into a computational type
$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows:
\[
  \begin{tikzcd}[sep=huge]
    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, "\IMode{\lambda(\phi)}"]
      \arrow[d, swap, "\IMode{\Pair{\ArrId{}}{\iota}}"]
      &
    \IMode{\Omega}\\
    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
  \end{tikzcd}
\]
where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
\begin{align*}
  \ADefine{
    \VToCType(\phi)(M,N)
  }{
    \forall M', N' \in \Values_\tau.\
    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
  }
\end{align*}

>>>>

  So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
  judgments in terms of the presheaf topos is that we do not need to deal with
  the complexities of quantifying over renamings, since this is implicit in the
  definition of the exponential of presheaves, where
  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-yoneda embedding:
  \begin{align*}
    \ADefine{
      B^A(\Upsilon)
    }{
      \Hom{\Sets^\SCtx}{
        \Coyoneda{\Upsilon}\times A
      }{
        B
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}
      (\Upsilon\inj\Upsilon')
      \Rightarrow
      {B(\Upsilon')}^{A(\Upsilon')}
    }
  \end{align*}

@@

%% In general, we must explain typehood for parametric refinements
%% $\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
%% Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
%% transformation
%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
%% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
%% is a unique $\Of{\psi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$ such that
%% the following diagram commutes, where
%% $\lambda(\chi)$ is the exponential transpose of $\chi$:
%% \[
%%   \IsValueType{\Upsilon}{\phi}{\tau}
%%   \triangleq
%%   \begin{tikzcd}[sep=huge]
%%     \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
%%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
%%       \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
%%     \IMode{\Omega}\\
%%     \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
%%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
%%   \end{tikzcd}
%% \]
%%
%% Equivalently, and perhaps more simply, we can express value typehood internally
%% in our topos as a presheaf:
%% \[
%%   \Define{
%%     \IsValueType{-}{\phi}{\tau}
%%   }{
%%     \sum_{\psi\in\Pow{\Values_\tau^2}}
%%     \prod_{M,N\in\Values_\tau}
%%     {\phi(M,N)}^{\psi(M,N)}
%%   }
%% \]

@@

Now, let us return briefly to Martin-L\"of's notion of type, which we call a
\emph{value type}. The minimal data of such a type's definition consists in a
partial equivalence relation on values,
$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
It is easy to turn such a relation into a computational type
$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows:
\[
  \begin{tikzcd}[sep=huge]
    \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, "\IMode{\lambda(\phi)}"]
      \arrow[d, swap, "\IMode{\Pair{\ArrId{}}{\iota}}"]
      &
    \IMode{\Omega}\\
    \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
  \end{tikzcd}
\]
where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
\begin{align*}
  \ADefine{
    \VToCType(\phi)(M,N)
  }{
    \forall M', N' \in \Values_\tau.\
    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
  }
\end{align*}
